---
title: "Hw5"
author: "Calvin Wong"
date: "4/8/2020"
output: html_document
---

```{r}
library(fpp2)
library(forecast)
library(seasonal)
library(gridExtra)
```

7.1
Consider the pigs series — the number of pigs slaughtered in Victoria each month.
a) Use the ses() function in R to find the optimal values of  
```{r}
sespigs <- ses(pigs, h = 4)
sespigs$model
```


b) Compute a 95% prediction interval for the first forecast using is the standard deviation of the residuals. Compare your interval with the interval produced by R.

```{r}
s <- sd(sespigs$residuals)
sespigs$mean[1] + 1.96*s
autoplot(sespigs) + autolayer(sespigs$fitted)
```


7.5
Data set books contains the daily sales of paperback and hardcover books at the same store. The task is to forecast the next four days’ sales for paperback and hardcover books.
a) Plot the series and discuss the main features of the data.

The sales of both paperback and hardcover have an increasing trend but there is no cyclic or seasonality factor seen.

```{r}
str(books)
autoplot(books)
```

b) Use the ses() function to forecast each series, and plot the forecasts.

```{r}
sespaperback <- ses(books[, "Paperback"], h = 4)
seshardcover <- ses(books[, "Hardcover"], h = 4)

autoplot(books[, "Paperback"], series = "Paperback") +
  autolayer(sespaperback, series = "Paperback") +
  autolayer(books[, "Hardcover"], series = "Hardcover") +
  autolayer(seshardcover, series = "Hardcover", PI = FALSE) +
  ylab("Books") +
  ggtitle("Quantity of paperback and hardcover books sold")
```

c) Compute the RMSE values for the training data in each case.

```{r}
ses_paperback <- sqrt(mean(sespaperback$residuals^2))
ses_hardcover <- sqrt(mean(seshardcover$residuals^2))

print(paste("Sqrt: ", ses_paperback))
print(paste("Mean: ", ses_hardcover))
```


7.6
We will continue with the daily sales of paperback and hardcover books in data set books.
a) Apply Holt’s linear method to the paperback and hardback series and compute four-day forecasts in each case.

```{r}
holtpaperback <- holt(books[, "Paperback"], h = 4)
holthardcover <- holt(books[, "Hardcover"], h = 4)

autoplot(books[, "Paperback"], series = "Paperback") +
  autolayer(holtpaperback, series = "Paperback") +
  autolayer(books[, "Hardcover"], series = "Hardcover") +
  autolayer(holtpaperback, series = "Hardcover") +
  ylab("Books") +
  ggtitle("Quantity of paperback and hardcover books sold")
```

b) Compare the RMSE measures of Holt’s method for the two series to those of simple exponential smoothing in the previous question. (Remember that Holt’s method is using one more parameter than SES.) Discuss the merits of the two forecasting methods for these data sets.

```{r}
s_paperback <- sqrt(mean(holtpaperback$residuals^2))
s_hardcover <- sqrt(mean(holthardcover$residuals^2))

print(paste("Sqrt: ", s_paperback))
print(paste("Mean: ", s_hardcover))
```

c) Compare the forecasts for the two series using both methods. Which do you think is best?

I think that the forecasts of hardcover sales were better than the ones of paperback sales. Because RMSE value is lower for hardcover sales. And because the forecasts of paperback sales couldn’t reflect the pattern in the data using Holt’s method.

d) Calculate a 95% prediction interval for the first forecast for each series, using the RMSE values and assuming normal errors. Compare your intervals with those produced using ses and holt.

```{r}
## 95% PI of paperback sales calculated by holt function
holtpaperback$upper[1, "95%"]
holtpaperback$lower[1, "95%"]

## 95% PI of paperback sales calculated by formula
holtpaperback$mean[1] + 1.96*s_paperback
holtpaperback$mean[1] - 1.96*s_paperback

## 95% PI of hardcover sales calculated by holt function
holthardcover$upper[1, "95%"]
holthardcover$lower[1, "95%"]

## 95% PI of hardcover sales calculated by formula
holthardcover$mean[1] + 1.96*s_hardcover
holthardcover$mean[1] - 1.96*s_hardcover
```

7.7
For this exercise use data set eggs, the price of a dozen eggs in the United States from 1900–1993. Experiment with the various options in the holt() function to see how much the forecasts change with damped trend, or with a Box-Cox transformation. Try to develop an intuition of what each argument is doing to the forecasts.
[Hint: use h=100 when calling holt() so you can clearly see the differences between the various options when plotting the forecasts.]

Which model gives the best RMSE?

```{r}
str(eggs)
autoplot(eggs)
```

```{r}
holteggs <- holt(eggs, h = 100)
autoplot(holteggs) +
  autolayer(holteggs$fitted)
```

```{r}
holtdamped_eggs <- holt(eggs, damped = TRUE, h = 100)
autoplot(holtdamped_eggs) +
  autolayer(holtdamped_eggs$fitted)
```

```{r}
holtBoxCox_eggs <- holt(eggs, 
                         lambda = BoxCox.lambda(eggs), 
                         h = 100)
autoplot(holtBoxCox_eggs) +
  autolayer(holtBoxCox_eggs$fitted)
```

```{r}
holtBoxCox_damped_eggs <- holt(
  eggs, 
  damped = TRUE,
  lambda = BoxCox.lambda(eggs),
  h = 100)
autoplot(holtBoxCox_damped_eggs) +
  autolayer(holtBoxCox_damped_eggs$fitted)
```

```{r}
holteggs <- sqrt(mean(holteggs$residuals^2))
holtdampeggs <- sqrt(mean(holtdamped_eggs$residuals^2))
holtboxcoxeggs <- sqrt(mean(holtBoxCox_eggs$residuals^2))
holtdampboxcoxeggs <- sqrt(mean(holtBoxCox_damped_eggs$residuals^2))

print(paste("Holt: ", holteggs))
print(paste("Holt Damp: ", holtdampeggs))
print(paste("Holt BoxCox: ", holtboxcoxeggs))
print(paste("Holt Damp Boxcox: ", holtdampboxcoxeggs))
```

7.8
Recall your retail time series data (from Exercise 3 in Section 2.10).

```{r}
retaildata <- readxl::read_excel("retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349909T"], frequency=12, start=c(1982,4))
autoplot(myts)
```

a) Why is multiplicative seasonality necessary for this series?



b) Apply Holt-Winters’ multiplicative method to the data. Experiment with making the trend damped.

```{r}
hw_retail <- hw(myts,
                     seasonal = "multiplicative")

hwd_retail <- hw(myts,
                      seasonal = "multiplicative",
                      damped = TRUE)

autoplot(hw_retail)
autoplot(hwd_retail)
```

c) Compare the RMSE of the one-step forecasts from the two methods. Which do you prefer?

When the RMSE values were compared, the damped model excelled. I also prefer the damped model because it will prohibit the limitless increase of sales forecast.

```{r}
error_hw_retail <- tsCV(myts, hw, h = 1, seasonal = "multiplicative")
error_hwd_retail <- tsCV(myts, hw, h = 1, seasonal = "multiplicative", damped = TRUE)

holtwmyts <- sqrt(mean(error_hw_retail^2, na.rm = TRUE))
holtwdmyts <- sqrt(mean(error_hwd_retail^2, na.rm = TRUE))

print(paste("Holt Winters': ", holtwmyts))
print(paste("Damped Holt Winters: ", holtwdmyts))
```

d) Check that the residuals from the best method look like white noise.

```{r}
checkresiduals(hwd_retail)
```

e) Now find the test set RMSE, while training the model to the end of 2010. Can you beat the seasonal naïve approach from Exercise 8 in Section 3.7?

```{r}
myts_retail_train <- window(myts, end = c(2010, 12))
myts_retail_test <- window(myts, start = 2011)


ets_hwd_retail_train <- hw(myts_retail_train, h = 36, seasonal = "multiplicative", damped = TRUE)

autoplot(ets_hwd_retail_train)
```

```{r}
accuracy(ets_hwd_retail_train, myts_retail_test)
```

7.9
For the same retail data, try an STL decomposition applied to the Box-Cox transformed series, followed by ETS on the seasonally adjusted data. How does that compare with your best previous forecasts on the test set?

```{r}
fc_stl_ets_retail_train <- myts_retail_train %>%
  stlm(
    #made stl model first
    s.window = 13,
    robust = TRUE,
    #designate that the seasonally adjusted data should be forecasted by ETS method.
    method = "ets",
    lambda = BoxCox.lambda(myts_retail_train)
  ) %>%
  #forecast using stl model
  forecast(
    h = 36,
    lambda = BoxCox.lambda(myts_retail_train)
    )
```

```{r}
autoplot(fc_stl_ets_retail_train)
```

```{r}
accuracy(fc_stl_ets_retail_train, myts_retail_test)
```

```{r}
fc_stl_ets_retail_train_without_tr <- 
  myts_retail_train %>%
    stlm(
      s.window = 13,
      robust = TRUE,
      method = "ets"
    ) %>%
    forecast(h = 36)

autoplot(fc_stl_ets_retail_train_without_tr)
```

```{r}
accuracy(fc_stl_ets_retail_train_without_tr, myts_retail_test)
```

